---
title: "Predict house prices: regression"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```



In a regression problem, we aim to predict the output of a continuous value, like a price or a probability. Contrast this with a classification problem, where we aim to predict a discrete label (for example, where a picture contains an apple or an orange).

This notebook builds a model to predict the median price of homes in a Boston suburb during the mid-1970s. To do this, we'll provide the model with some data points about the suburb, such as the crime rate and the local property tax rate.


```{r}
library(keras)
```


### The Boston Housing Prices dataset

[This dataset](https://www.cs.toronto.edu/~delve/data/boston/bostonDetail.html) is accessible directly from keras.

```{r}
boston_housing <- dataset_boston_housing()

train_data <- boston_housing$train$x
train_labels <- boston_housing$train$y
test_data <- boston_housing$test$x
test_labels <- boston_housing$test$y
```

