---
title: "Classify movie reviews: binary classification"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, cache = TRUE)
```

This notebook classifies movie reviews as positive or negative using the text of the review. This is an example of binary — or two-class — classification, an important and widely applicable kind of machine learning problem.

We'll use the [IMDB dataset](https://tensorflow.rstudio.com/keras/reference/dataset_imdb.html) that contains the text of 50,000 movie reviews from the [Internet Movie Database](https://www.imdb.com/). These are split into 25,000 reviews for training and 25,000 reviews for testing. The training and testing sets are balanced, meaning they contain an equal number of positive and negative reviews.

Let's start and load Keras, as well as a few other required libraries.


```{r}
library(keras)
library(dplyr)
library(ggplot2)
library(purrr)
```


### Download the IMDB dataset

The IMDB dataset comes packaged with Keras. It has already been preprocessed such that the reviews (sequences of words) have been converted to sequences of integers, where each integer represents a specific word in a dictionary.

The following code downloads the IMDB dataset to your machine (or uses a cached copy if you've already downloaded it):

```{r}
imdb <- dataset_imdb(num_words = 10000)

train_data <- imdb$train[[1]]
train_labels <- imdb$train[[2]]

test_data <- imdb$test[[1]]
test_labels <- imdb$test[[2]]
```


The argument num_words = 10000 keeps the top 10,000 most frequently occurring words in the training data. The rare words are discarded to keep the size of the data manageable.

Conveniently, the dataset comes with an index mapping words to integers, which has to be downloaded separately:

```{r}
word_index <- dataset_imdb_word_index()
```


### Explore the data

Let's take a moment to understand the format of the data. The dataset comes preprocessed: each example is an array of integers representing the words of the movie review. Each label is an integer value of either 0 or 1, where 0 is a negative review, and 1 is a positive review.

```{r}
paste0("Training entries: ", length(train_data), ", labels: ", length(train_labels))
```


The texts of the reviews have been converted to integers, where each integer represents a specific word in a dictionary. Here's what the first review looks like:

```{r}
train_data[[1]]
```

Movie reviews may be different lengths. The below code shows the number of words in the first and second reviews. Since inputs to a neural network must be the same length, we'll need to resolve this later.

```{r}
length(train_data[[1]])
length(train_data[[2]])
```



##### Convert the integers back to words

It may be useful to know how to convert integers back to text.
We already have the `word_index` we downloaded above — a list with words as keys and integers as values. If we create a data.frame from it, we can conveniently use it in both directions.

```
Is this ok? I was fiddling around creating a reversed list and running into string to integer  conversion problems and the like and then thought maybe this is easiest?
```

```{r}
word_index_df <- data.frame(
  word = names(word_index),
  idx = unlist(word_index, use.names = FALSE),
  stringsAsFactors = FALSE
)

# The first indices are reserved  
word_index_df <- word_index_df %>% mutate(idx = idx + 3)
word_index_df <- word_index_df %>%
  add_row(word = "<PAD>", idx = 0)%>%
  add_row(word = "<START>", idx = 1)%>%
  add_row(word = "<UNK>", idx = 2)%>%
  add_row(word = "<UNUSED>", idx = 3)

word_index_df <- word_index_df %>% arrange(idx)

decode_review <- function(text){
  paste(map(text, function(number) word_index_df %>%
              filter(idx == number) %>%
              select(word) %>% 
              pull()),
        collapse = " ")
}
```


Now we can use the decode_review function to display the text for the first review:

```{r}
decode_review(train_data[[1]])
```




### Prepare the data

The reviews — the arrays of integers — must be converted to tensors before fed into the neural network. This conversion can be done a couple of ways:

 - One-hot-encode the arrays to convert them into vectors of 0s and 1s. For example, the sequence [3, 5] would become a 10,000-dimensional vector that is all zeros except for indices 3 and 5, which are ones. Then, make this the first layer in our network — a `dense` layer — that can handle floating point vector data. This approach is memory intensive, though, requiring a `num_words * num_reviews` size matrix.

 - Alternatively, we can pad the arrays so they all have the same length, then create an integer tensor of shape `num_examples * max_length`. We can use an embedding layer capable of handling this shape as the first layer in our network.

In this tutorial, we will use the second approach.

Since the movie reviews must be the same length, we will use the [pad_sequences](https://tensorflow.rstudio.com/keras/reference/pad_sequences.html) function to standardize the lengths:

```{r}
train_data <- pad_sequences(
  train_data,
  value = word_index_df %>% filter(word == "<PAD>") %>% select(idx) %>% pull(),
  padding = "post",
  maxlen = 256
)

test_data <- pad_sequences(
  test_data,
  value = word_index_df %>% filter(word == "<PAD>") %>% select(idx) %>% pull(),
  padding = "post",
  maxlen = 256
)
```


Let's look at the length of the examples now:

```{r}
length(train_data[1, ])
length(train_data[2, ])
```

And inspect the (now padded) first review:

```{r}
train_data[1, ]
```



### Build the model

The neural network is created by stacking layers — this requires two main architectural decisions:

- How many layers to use in the model?
- How many hidden units to use for each layer?

In this example, the input data consists of an array of word-indices. The labels to predict are either 0 or 1. Let's build a model for this problem:

```{r}
# input shape is the vocabulary count used for the movie reviews (10,000 words)
vocab_size <- 10000

model <- keras_model_sequential()
model %>% 
  layer_embedding(input_dim = vocab_size, output_dim = 16) %>%
  layer_global_average_pooling_1d() %>%
  layer_dense(units = 16, activation = "relu") %>%
  layer_dense(units = 1, activation = "sigmoid")

model %>% summary()
```




The layers are stacked sequentially to build the classifier:

 - The first layer is an `embedding` layer. This layer takes the integer-encoded vocabulary and looks up the embedding vector for each word-index. These vectors are learned as the model trains. The vectors add a dimension to the output array. The resulting dimensions are: (`batch, sequence, embedding`).
 - Next, a `global_average_pooling_1d` layer returns a fixed-length output vector for each example by averaging over the sequence dimension. This allows the model to handle input of variable length, in the simplest way possible.
 - This fixed-length output vector is piped through a fully-connected (`dense`) layer with 16 hidden units.
 - The last layer is densely connected with a single output node. Using the `sigmoid` activation function, this value is a float between 0 and 1, representing a probability, or confidence level.



#### Hidden units

The above model has two intermediate or "hidden" layers, between the input and output. The number of outputs (units, nodes, or neurons) is the dimension of the representational space for the layer. In other words, the amount of freedom the network is allowed when learning an internal representation.

If a model has more hidden units (a higher-dimensional representation space), and/or more layers, then the network can learn more complex representations. However, it makes the network more computationally expensive and may lead to learning unwanted patterns — patterns that improve performance on training data but not on the test data. This is called overfitting, and we'll explore it later.


#### Loss function and optimizer

A model needs a loss function and an optimizer for training. Since this is a binary classification problem and the model outputs a probability (a single-unit layer with a sigmoid activation), we'll use the `binary_crossentropy` loss function.

This isn't the only choice for a loss function, you could, for instance, choose `mean_squared_error`. But, generally, binary_crossentropy is better for dealing with probabilities — it measures the "distance" between probability distributions, or in our case, between the ground-truth distribution and the predictions.

Later, when we are exploring regression problems (say, to predict the price of a house), we will see how to use another loss function called mean squared error.

Now, configure the model to use an optimizer and a loss function:

```{r}
model %>% compile(
  optimizer = 'adam',
  loss = 'binary_crossentropy',
  metrics = list('accuracy')
)
```


### Create a validation set

When training, we want to check the accuracy of the model on data it hasn't seen before. Create a validation set by setting apart 10,000 examples from the original training data. (Why not use the testing set now? Our goal is to develop and tune our model using only the training data, then use the test data just once to evaluate our accuracy).


```{r}
x_val <- train_data[1:10000, ]
partial_x_train <- train_data[10001:nrow(train_data), ]

y_val <- train_labels[1:10000]
partial_y_train <- train_labels[10001:length(train_labels)]
```


### Train the model

Train the model for 20 epochs in mini-batches of 512 samples. This is 20 iterations over all samples in the x_train and y_train tensors. While training, monitor the model's loss and accuracy on the 10,000 samples from the validation set:

```{r}
history <- model %>% fit(
  partial_x_train,
  partial_y_train,
  epochs = 40,
  batch_size = 512,
  validation_data = list(x_val, y_val),
  verbose=1
)
```



### Evaluate the model

And let's see how the model performs. Two values will be returned. Loss (a number which represents our error, lower values are better), and accuracy.

```{r}
results <- model %>% evaluate(test_data, test_labels)
results
```


This fairly naive approach achieves an accuracy of about 87%. With more advanced approaches, the model should get closer to 95%.


### Create a graph of accuracy and loss over time

`model$fit` returns a `keras_training_history` object whose `metrics` slot contains loss and metrics values recorded during training.
You can conveniently plot the loss and metrics curves like so:


```{r}
plot(history)
```


The evolution of loss and metrics can also be seen during training in the RStudio Viewer pane.

Notice the training loss decreases with each epoch and the training accuracy increases with each epoch. This is expected when using gradient descent optimization — it should minimize the desired quantity on every iteration.

This isn't the case for the validation loss and accuracy — they seem to peak after about twenty epochs. This is an example of overfitting: the model performs better on the training data than it does on data it has never seen before. After this point, the model over-optimizes and learns representations specific to the training data that do not generalize to test data.

For this particular case, we could prevent overfitting by simply stopping the training after twenty or so epochs. Later, you'll see how to do this automatically with a callback.
